---
title: "Hohmeier_Project4"
author: "Kaitlyn Hohmeier"
date: "11/9/2020"
output: pdf_document
---
  <style type="text/css">
  
  body{ /* Normal  */
      font-size: 16px;
  }
td {  /* Table  */
    font-size: 8px;
}
h1.title {
  font-size: 38px;
  color: DarkRed;
}
h1 { /* Header 1 */
    font-size: 28px;
  color: DarkBlue;
}
h2 { /* Header 2 */
    font-size: 22px;
  color: DarkBlue;
}
h3 { /* Header 3 */
    font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: DarkBlue;
}
code.r{ /* Code block */
    font-size: 16px;
}
pre { /* Code block - determines code spacing between lines */
    font-size: 16px;
}
</style>

```{r}
library(MASS);
library(ISLR);
library(car);
library(tidyverse);
library(class);
```


```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE, comment = NA, message = FALSE)
```

```{r, echo = F}
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=80),tidy=TRUE)
```
## R Markdown ##

```{r}
# Initial data read-in, without "cleaning up"

init <- read.csv("C://Users//Kaitlyn//Documents//Documents - Old Laptop//College Work//3rd year Junior//Semester 1//STT 450//Project//US Census Data Sets//acs2015_census_tract_data.csv")
```

```{r}
# Cleaning up the data
Income <- na.omit(init)
```

### Section 5: Classification ###

### Step 1: Train on Full, Test on Full ###
### Part a. Logistic Regression###

```{r}
# First, we need to create a new factor variable from Income per Capita, based on national poverty level of $12,084 for an individual
new_Income = Income
levels <- ifelse(new_Income$IncPerCap <= 12082, "low", "high")
new_Income$Income_Levels <- levels
table(new_Income$Income_Levels)

binary_levels <- ifelse(new_Income$Income_Levels=="high", 1, 0)
new_Income$Income_Level_Binary <- binary_levels
```

```{r}
start.time <- Sys.time()
glm.fit2=glm(Income_Level_Binary~Men + Women + Hispanic + Black + Native + Poverty + ChildPov + Professional + Service + Construction,data=new_Income,family=binomial)
end.time <- Sys.time()
time.taken <- end.time - start.time
time.taken
summary(glm.fit2)
```

```{r}
glm.probs2=predict(glm.fit2,type="response")
glm.probs2[1:100]
```

```{r}
# Run contrasts
# contrasts(new_Income$factor(Income_Level_Binary))
```


```{r}
# Create confusion matrix
glm.pred=rep("low",72727)
glm.pred[glm.probs2>=.5]="high"
# Prediction accuracy
acc_log = mean(glm.pred==new_Income$Income_Levels)
acc_log
```
```{r}
Confusion = table(glm.pred, new_Income$Income_Levels)
Confusion
Sensitivity_log = Confusion[1,1]/(Confusion[1,1]+Confusion[2,1]); Sensitivity_log
Specificity_log = Confusion[2,2]/(Confusion[2,2]+Confusion[1,2]); Specificity_log
```
###Part b. LDA###
```{r}
# Perform LDA
start.time <- Sys.time()
lda.fit=MASS::lda(Income_Level_Binary~Men + Women + Hispanic + Black + Native + Poverty + ChildPov + Professional + Service + Construction,data=new_Income)
end.time <- Sys.time()
time.taken <- end.time - start.time
time.taken
lda.fit
```
```{r}
# Plot the LDA analysis that was run
plot(lda.fit)
```

```{r}
# Use LDA to make predictions
lda.pred=predict(lda.fit, new_Income)
names(lda.pred)
lda.class=lda.pred$class;
head(lda.class)
```

```{r}
# for LDA prediction posterior for classes
head(lda.pred$posterior)
head(lda.pred$x)
```


```{r}
# Confusion matrix for LDA
confusion <- table(lda.class,new_Income$Income_Level_Binary)
rownames(confusion) <- c("low","high")
colnames(confusion) <- c("low","high")
confusion
Sensitivity_lda = confusion[2,2]/(confusion[1,2]+confusion[2,2]); Sensitivity_lda
Specificity_lda = confusion[1,1]/(confusion[1,1]+confusion[2,1]); Specificity_lda

```

```{r}
# Predictor accuracy
acc_lda = mean(lda.class==new_Income$Income_Level_Binary)
acc_lda
```


###Part c. QDA ###

```{r}
# Perform QDA
start.time <- Sys.time()
qda.fit=MASS::qda(Income_Level_Binary~Men + Women + Hispanic + Black + Native + Poverty + ChildPov + Professional + Service + Construction,data=new_Income)
end.time <- Sys.time()
time.taken <- end.time - start.time
time.taken
qda.fit
```

```{r}
# Create confusion matrix
qda.class=predict(qda.fit,new_Income)$class
confusion_qda <- table(qda.class,new_Income$Income_Level_Binary)
rownames(confusion_qda) <- c("low","high")
colnames(confusion_qda) <- c("low","high")
confusion_qda
Sensitivity_qda= confusion_qda[2,2]/(confusion_qda[1,2]+confusion_qda[2,2]); Sensitivity_qda
Specificity_qda= confusion_qda[1,1]/(confusion_qda[1,1]+confusion_qda[2,1]); Specificity_qda
```

```{r}
acc_qda = mean(qda.class==new_Income$Income_Level_Binary)
acc_qda
```

###Part d. KNN ###
```{r}
# Create test and training data sets. In this case, we are testing on full, training on full
train.X=cbind(new_Income$Men, new_Income$Women, new_Income$Hispanic, new_Income$Black, new_Income$Native, new_Income$Poverty, new_Income$ChildPov, new_Income$Professional, new_Income$Service, new_Income$Construction)
test.X=cbind(new_Income$Men, new_Income$Women, new_Income$Hispanic, new_Income$Black, new_Income$Native, new_Income$Poverty, new_Income$ChildPov, new_Income$Professional, new_Income$Service, new_Income$Construction)
train.Income=new_Income$Income_Levels
```


```{r}
# Run KNN k = 1
start.time <- Sys.time()
set.seed(1)
knn.pred=knn(train.X,test.X,train.Income,k=1)
end.time <- Sys.time()
time.taken <- end.time - start.time
time.taken
head(knn.pred)
```

```{r}
# Create Confusion matrix, k = 1
confusion_k1 <- table(knn.pred,new_Income$Income_Levels)
confusion_k1
Sensitivity_k1= confusion_k1[1,1]/(confusion_k1[1,1]+confusion_k1[2,1]); Sensitivity_k1
Specificity_k1= confusion_k1[2,2]/(confusion_k1[2,2]+confusion_k1[1,2]); Specificity_k1
```

```{r}
# Prediction accuracy k = 1
acc_k1 = mean(knn.pred==new_Income$Income_Levels)
acc_k1
```

```{r}
# Run KNN k = 2
start.time <- Sys.time()
knn.pred2=knn(train.X,test.X,train.Income,k=2)
end.time <- Sys.time()
time.taken <- end.time - start.time
time.taken
head(knn.pred2)
```

```{r}
# Create Confusion matrix, k = 2
confusion_k2 <- table(knn.pred2,new_Income$Income_Levels)
confusion_k2
Sensitivity_k2= confusion_k2[1,1]/(confusion_k2[1,1]+confusion_k2[2,1]); Sensitivity_k2
Specificity_k2= confusion_k2[2,2]/(confusion_k2[2,2]+confusion_k2[1,2]); Specificity_k2
```

```{r}
# Prediction accuracy, k = 2
acc_k2 = mean(knn.pred2==new_Income$Income_Levels)
acc_k2
```
```{r}
# Run KNN k = 3
start.time <- Sys.time()
knn.pred3=knn(train.X,test.X,train.Income,k=3)
end.time <- Sys.time()
time.taken <- end.time - start.time
time.taken
head(knn.pred3)
```

```{r}
# Create Confusion matrix, k = 3
confusion_k3 <- table(knn.pred3,new_Income$Income_Levels)
confusion_k3
Sensitivity_k3= confusion_k3[1,1]/(confusion_k3[1,1]+confusion_k3[2,1]); Sensitivity_k3
Specificity_k3= confusion_k3[2,2]/(confusion_k3[2,2]+confusion_k3[1,2]); Specificity_k3
```

```{r}
# Prediction accuracy, k = 3
acc_k3 = mean(knn.pred3==new_Income$Income_Levels)
acc_k3
```

```{r}
# Run KNN k = 4
start.time <- Sys.time()
knn.pred4=knn(train.X,test.X,train.Income,k=4)
end.time <- Sys.time()
time.taken <- end.time - start.time
time.taken
head(knn.pred4)
```

```{r}
# Create Confusion matrix, k = 4
confusion_k4 <- table(knn.pred4,new_Income$Income_Levels)
confusion_k4
Sensitivity_k4= confusion_k4[1,1]/(confusion_k4[1,1]+confusion_k4[2,1]); Sensitivity_k4
Specificity_k4= confusion_k4[2,2]/(confusion_k4[2,2]+confusion_k4[1,2]); Specificity_k4
```

```{r}
# Prediction accuracy, k = 4
acc_k4 = mean(knn.pred4==new_Income$Income_Levels)
acc_k4
```

```{r}
# Run KNN k = 5
start.time <- Sys.time()
knn.pred5=knn(train.X,test.X,train.Income,k=5)
end.time <- Sys.time()
time.taken <- end.time - start.time
time.taken
head(knn.pred5)
```

```{r}
# Create Confusion matrix, k = 5
confusion_k5 <- table(knn.pred5,new_Income$Income_Levels)
confusion_k5
Sensitivity_k5= confusion_k5[1,1]/(confusion_k5[1,1]+confusion_k5[2,1]); Sensitivity_k5
Specificity_k6= confusion_k5[2,2]/(confusion_k5[2,2]+confusion_k5[1,2]); Specificity_k6
```

```{r}
# Prediction accuracy, k = 5
acc_k5 = mean(knn.pred5==new_Income$Income_Levels)
acc_k5
```

```{r}
# Run KNN k = 6
start.time <- Sys.time()
knn.pred6=knn(train.X,test.X,train.Income,k=6)
end.time <- Sys.time()
time.taken <- end.time - start.time
time.taken
head(knn.pred6)
```

```{r}
# Create Confusion matrix, k = 6
confusion_k6 <- table(knn.pred6,new_Income$Income_Levels)
confusion_k6
Sensitivity_k6= confusion_k6[1,1]/(confusion_k6[1,1]+confusion_k6[2,1]); Sensitivity_k6
Specificity_k6= confusion_k6[2,2]/(confusion_k6[2,2]+confusion_k6[1,2]); Specificity_k6
```

```{r}
# Prediction accuracy, k = 6
acc_k6 = mean(knn.pred6==new_Income$Income_Levels)
acc_k6
```

```{r}
# Run KNN k = 7
start.time <- Sys.time()
knn.pred7=knn(train.X,test.X,train.Income,k=7)
end.time <- Sys.time()
time.taken <- end.time - start.time
time.taken
head(knn.pred7)
```

```{r}
# Create Confusion matrix, k = 7
confusion_k7 <- table(knn.pred7,new_Income$Income_Levels)
confusion_k7
Sensitivity_k7= confusion_k7[1,1]/(confusion_k7[1,1]+confusion_k7[2,1]); Sensitivity_k7
Specificity_k7= confusion_k7[2,2]/(confusion_k7[2,2]+confusion_k7[1,2]); Specificity_k7
```

```{r}
# Prediction accuracy, k = 7
acc_k7 = mean(knn.pred7==new_Income$Income_Levels)
acc_k7
```

### Step 2: Train on 50%, Test on 50% ###
```{r}
# Set up training and test sets. First 50% of data is training set; remaining 50% is testing set
test=1:36364
training.x = new_Income[test,]
testing.x = new_Income[-test,]

training.y = new_Income$Income_Levels[test]
testing.y = new_Income$Income_Levels[-test]
```

### Part a. Logistic Regression ###

```{r}
# Perform logistic regression on training set
start.time <- Sys.time()
glm.fit=glm(Income_Level_Binary~Men + Women + Hispanic + Black + Native + Poverty + ChildPov + Professional + Service + Construction, data=new_Income, family=binomial, subset=test)
end.time <- Sys.time()
time.taken <- end.time - start.time
time.taken
summary(glm.fit)
```
```{r}
# Use glm models to predict for the test set
glm.probs=predict(glm.fit,new_Income[-test,],type="response")
glm.pred=rep("low",36363)
glm.pred[glm.probs>.5]="high"

# Create confusion matrix
Confusion_log <- table(glm.pred,testing.y)
Confusion_log
Sensitivity_log2= Confusion_log[1,1]/(Confusion_log[1,1]+Confusion_log[2,1]); Sensitivity_log2
Specificity_log2= Confusion_log[2,2]/(Confusion_log[2,2]+Confusion_log[1,2]); Specificity_log2
```
```{r}
# prediction accuracy
acc_log2 = mean(glm.pred==testing.y)
acc_log2
```

###Part b. LDA ###

```{r}
training.y = new_Income$Income_Level_Binary[test]
testing.y = new_Income$Income_Level_Binary[-test]
```


```{r}
# Perform LDA
start.time <- Sys.time()
lda.fit2=MASS::lda(Income_Level_Binary~Men + Women + Hispanic + Black + Native + Poverty + ChildPov + Professional + Service + Construction,data=new_Income,subset=test)
end.time <- Sys.time()
time.taken <- end.time - start.time
time.taken
lda.fit2
```
```{r}
# Plot the LDA analysis that was run
plot(lda.fit2)
```

```{r}
# Use LDA to make predictions
lda.pred2=predict(lda.fit2, testing.x)
names(lda.pred2)
lda.class2=lda.pred2$class;
head(lda.class2)
```

```{r}
# for LDA prediction posterior for classes
head(lda.pred2$posterior)
head(lda.pred2$x)
```


```{r}
# Confusion matrix for LDA
confusion <- table(lda.class2,testing.y)
rownames(confusion) <- c("low","high")
colnames(confusion) <- c("low","high")
confusion
Sensitivity_lda2= confusion[2,2]/(confusion[1,2]+confusion[2,2]); Sensitivity_lda2
Specificity_lda2= confusion[1,1]/(confusion[1,1]+confusion[2,1]); Specificity_lda2
```

```{r}
# Predictor accuracy
acc_lda2 = mean(lda.class2==testing.y)
acc_lda2
```

###Part c. QDA ###

```{r}
# Perform QDA
start.time <- Sys.time()
qda.fit2=MASS::qda(Income_Level_Binary~Men + Women + Hispanic + Black + Native + Poverty + ChildPov + Professional + Service + Construction,data=new_Income,subset=test)
end.time <- Sys.time()
time.taken <- end.time - start.time
time.taken
qda.fit2
```
```{r}
# Create confusion matrix
qda.class2=predict(qda.fit2,testing.x)$class
confusion_qda2 <- table(qda.class2,testing.y)
rownames(confusion_qda2) <- c("low","high")
colnames(confusion_qda2) <- c("low","high")
confusion_qda2
Sensitivity_qda2= confusion_qda2[2,2]/(confusion_qda2[1,2]+confusion_qda2[2,2]); Sensitivity_qda2
Specificity_qda2= confusion_qda2[1,1]/(confusion_qda2[1,1]+confusion_qda2[2,1]); Specificity_qda2
```

```{r}
# prediction accuracy
acc_qda2 = mean(qda.class2==testing.y)
acc_qda2
```

###Part d. KNN ###

```{r}
# Create test and training data sets. In this case, we are training on the first 50% and testing on the last 50%
# Note: an observation was removed from the training set in order to make the training and testing set sizes equal. Otherwise, KNN would not run.
train.X=cbind(new_Income$Men, new_Income$Women, new_Income$Hispanic, new_Income$Black, new_Income$Native, new_Income$Poverty, new_Income$ChildPov, new_Income$Professional, new_Income$Service, new_Income$Construction)[test,]
test.X=cbind(new_Income$Men, new_Income$Women, new_Income$Hispanic, new_Income$Black, new_Income$Native, new_Income$Poverty, new_Income$ChildPov, new_Income$Professional, new_Income$Service, new_Income$Construction)[-test,]
train.Y=new_Income$Income_Levels[test]
test.Y=new_Income$Income_Levels[-test]
```

```{r}
# Run KNN, k = 1
# set.seed(1)
start.time <- Sys.time()
knn.pred11=knn(train.X,test.X,train.Y,k=1)
end.time <- Sys.time()
time.taken <- end.time - start.time
time.taken
head(knn.pred11)
```
```{r}
# Create Confusion matrix, k = 1
confusion_k11 <- table(knn.pred11,test.Y)
confusion_k11
Sensitivity_k11= confusion_k11[1,1]/(confusion_k11[1,1]+confusion_k11[2,1]); Sensitivity_k11
Specificity_k11= confusion_k11[2,2]/(confusion_k11[2,2]+confusion_k11[1,2]); Specificity_k11
```

```{r}
# Prediction accuracy k = 1
acc_k11 = mean(knn.pred11==test.Y)
acc_k11
```
```{r}
# Run KNN, k = 2
# set.seed(1)
start.time <- Sys.time()
knn.pred12=knn(train.X,test.X,train.Y,k=2)
end.time <- Sys.time()
time.taken <- end.time - start.time
time.taken
head(knn.pred12)
```

```{r}
# Create Confusion matrix, k = 2
confusion_k12 <- table(knn.pred12,test.Y)
confusion_k12
Sensitivity_k12= confusion_k12[1,1]/(confusion_k12[1,1]+confusion_k12[2,1]); Sensitivity_k12
Specificity_k12= confusion_k12[2,2]/(confusion_k12[2,2]+confusion_k12[1,2]); Specificity_k12
```

```{r}
# Prediction accuracy k = 2
acc_k12 = mean(knn.pred12==test.Y)
acc_k12
```

```{r}
# Run KNN, k = 3
# set.seed(1)
start.time <- Sys.time()
knn.pred13=knn(train.X,test.X,train.Y,k=3)
end.time <- Sys.time()
time.taken <- end.time - start.time
time.taken
head(knn.pred13)
```

```{r}
# Create Confusion matrix, k = 3
confusion_k13 <- table(knn.pred13,test.Y)
confusion_k13
Sensitivity_k13= confusion_k13[1,1]/(confusion_k13[1,1]+confusion_k13[2,1]); Sensitivity_k13
Specificity_k13= confusion_k13[2,2]/(confusion_k13[2,2]+confusion_k13[1,2]); Specificity_k13
```

```{r}
# Prediction accuracy k = 3
acc_k13 = mean(knn.pred13==test.Y)
acc_k13
```
```{r}
# Run KNN, k = 4
# set.seed(1)
start.time <- Sys.time()
knn.pred14=knn(train.X,test.X,train.Y,k=4)
end.time <- Sys.time()
time.taken <- end.time - start.time
time.taken
head(knn.pred14)
```

```{r}
# Create Confusion matrix, k = 4
confusion_k14 <- table(knn.pred14,test.Y)
confusion_k14
Sensitivity_k14= confusion_k14[1,1]/(confusion_k14[1,1]+confusion_k14[2,1]); Sensitivity_k14
Specificity_k14= confusion_k14[2,2]/(confusion_k14[2,2]+confusion_k14[1,2]); Specificity_k14
```

```{r}
# Prediction accuracy k = 4
acc_k14 = mean(knn.pred14==test.Y)
acc_k14
```
```{r}
# Run KNN, k = 5
# set.seed(1)
start.time <- Sys.time()
knn.pred15=knn(train.X,test.X,train.Y,k=5)
end.time <- Sys.time()
time.taken <- end.time - start.time
time.taken
head(knn.pred15)
```

```{r}
# Create Confusion matrix, k = 5
confusion_k15 <- table(knn.pred15,test.Y)
confusion_k15
Sensitivity_k15= confusion_k15[1,1]/(confusion_k15[1,1]+confusion_k15[2,1]); Sensitivity_k15
Specificity_k15= confusion_k15[2,2]/(confusion_k15[2,2]+confusion_k15[1,2]); Specificity_k15
```

```{r}
# Prediction accuracy k = 5
acc_k15 = mean(knn.pred15==test.Y)
acc_k15
```
```{r}
# Run KNN, k = 6
# set.seed(1)
start.time <- Sys.time()
knn.pred16=knn(train.X,test.X,train.Y,k=6)
end.time <- Sys.time()
time.taken <- end.time - start.time
time.taken
head(knn.pred16)
```

```{r}
# Create Confusion matrix, k = 6
confusion_k16 <- table(knn.pred16,test.Y)
confusion_k16
Sensitivity_k16= confusion_k16[1,1]/(confusion_k16[1,1]+confusion_k16[2,1]); Sensitivity_k16
Specificity_k16= confusion_k16[2,2]/(confusion_k16[2,2]+confusion_k16[1,2]); Specificity_k16
```

```{r}
# Prediction accuracy k = 6
acc_k16 = mean(knn.pred16==test.Y)
acc_k16
```
```{r}
# Run KNN, k = 7
# set.seed(1)
start.time <- Sys.time()
knn.pred17=knn(train.X,test.X,train.Y,k=7)
end.time <- Sys.time()
time.taken <- end.time - start.time
time.taken
head(knn.pred16)
```

```{r}
# Create Confusion matrix, k = 7
confusion_k17 <- table(knn.pred17,test.Y)
confusion_k17
Sensitivity_k17= confusion_k17[1,1]/(confusion_k17[1,1]+confusion_k17[2,1]); Sensitivity_k17
Specificity_k17= confusion_k17[2,2]/(confusion_k17[2,2]+confusion_k17[1,2]); Specificity_k17
```

```{r}
# Prediction accuracy k = 7
acc_k17 = mean(knn.pred17==test.Y)
acc_k17
```

###Part 3. Graphical Comparisons ###
```{r}
# Prepare results for graphing in barchart
graphing_data_full = cbind(acc_log, Sensitivity_log, Specificity_log, acc_lda, Sensitivity_lda, Specificity_lda, acc_qda, Sensitivity_qda, Specificity_qda, acc_k3, Sensitivity_k3, Specificity_k3)
graphing_data_full
results_full <- matrix(graphing_data_full, nrow=4,ncol=3,byrow = T)
colnames(results_full) <- c("Accuracy", "Sensitivity", "Specificity")
rownames(results_full) <- c("Logistic Reg", "LDA", "QDA", "KNN")
graphing_data_full
results_full
```
```{r}
barplot(results_full, 
        col=c("cornflowerblue","bisque4","brown","burlywood"), 
        border="white", 
        font.axis=2, 
        beside=T, 
        #legend=rownames(results_full), 
        xlab="Logistic Reg, LDA, QDA, and KNN k = 3, Full Data", 
        font.lab=2)
legend("topright", legend= rownames(results_full), fill =c("cornflowerblue","bisque4","brown","burlywood"), box.lty=0)
```
```{r}
# Prepare results for graphing in barchart
graphing_data_half = cbind(acc_log2, Sensitivity_log2, Specificity_log2, acc_lda2, Sensitivity_lda2, Specificity_lda2, acc_qda2, Sensitivity_qda2, Specificity_qda2, acc_k13, Sensitivity_k13, Specificity_k13)
graphing_data_half
results_half <- matrix(graphing_data_half, nrow=4,ncol=3,byrow = T)
colnames(results_half) <- c("Accuracy", "Sensitivity", "Specificity")
rownames(results_half) <- c("Logistic Reg", "LDA", "QDA", "KNN")
graphing_data_half
results_half
```
```{r}
barplot(results_half, 
        col=c("cornflowerblue","bisque4","brown","burlywood"), 
        border="white", 
        font.axis=2, 
        beside=T, 
        #legend=rownames(results_half), 
        xlab="Logistic Reg, LDA, QDA, and KNN k = 3, 50% Data", 
        font.lab=2)
legend("topright", legend= rownames(results_half), fill =c("cornflowerblue","bisque4","brown","burlywood"), box.lty=0)
```














