---
title: "Hohmeier_Project3"
author: "Kaitlyn Hohmeier"
date: "10/5/2020"
output: pdf_document
---
  <style type="text/css">
  
  body{ /* Normal  */
      font-size: 16px;
  }
td {  /* Table  */
    font-size: 8px;
}
h1.title {
  font-size: 38px;
  color: DarkRed;
}
h1 { /* Header 1 */
    font-size: 28px;
  color: DarkBlue;
}
h2 { /* Header 2 */
    font-size: 22px;
  color: DarkBlue;
}
h3 { /* Header 3 */
    font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: DarkBlue;
}
code.r{ /* Code block */
    font-size: 16px;
}
pre { /* Code block - determines code spacing between lines */
    font-size: 16px;
}
</style>


```{r}
library(leaps)
library(formatR)
library(knitr)
```


```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE, comment = NA, message = FALSE)
```

```{r, echo = F}
opts_chunk$set(tidy.opts=list(width.cutoff=80),tidy=TRUE)
```
## R Markdown ##

```{r}
# Initial data read-in, without "cleaning up"

init <- read.csv("C://Users//Kaitlyn//Documents//Documents - Old Laptop//College Work//3rd year Junior//Semester 1//STT 450//Project//US Census Data Sets//acs2015_census_tract_data.csv")
```

```{r}
# Cleaning up the data
Income <- na.omit(init)
```

### Section 4: Classification ###

### Step 1: Best Subset Selection ###

```{r}
subset_data = subset(Income, select = -c(State,County,CensusTract,TotalPop,IncomeErr,IncomePerCapErr, Drive, Carpool, Transit, Walk, OtherTransp, WorkAtHome, MeanCommute, Employed, PrivateWork,PublicWork,SelfEmployed, FamilyWork,Unemployment, Citizen))
regfit.full=regsubsets(Income~.,subset_data,nvmax=16)
reg.summary = summary(regfit.full)
reg.summary
```

```{r}
#Create plots for RSS, Adjuted R-squared, Mallow's Cp, and BIC - Best Subset
par(mfrow=c(2,2))
#Plot RSS
plot(reg.summary$rss,xlab="Number of Variables",ylab="RSS",type="l")

#Plot Adjusted R-squared
plot(reg.summary$adjr2,xlab="Number of Variables",ylab="Adjusted RSq",type="l")
which.max(reg.summary$adjr2)
points(15,reg.summary$adjr2[15], col="orange",cex=2,pch=20)

#Plot Mallow's Cp
plot(reg.summary$cp,xlab="Number of Variables",ylab="Cp",type='l')
which.min(reg.summary$cp)
points(14,reg.summary$cp[14],col="orange",cex=2,pch=20)

#Plot BIC
plot(reg.summary$bic,xlab="Number of Variables",ylab="BIC",type='l')
which.min(reg.summary$bic)
points(12,reg.summary$bic[12],col="orange",cex=2,pch=20)
```

```{r}
#Create best model based on results of best subset selection - log (Cp)
lm.fit1a=lm(log1p(Income)~log1p(Men) + log1p(Women) + log1p(Hispanic) + log1p(White) + log1p(Black) + log1p(Native) + log1p(Asian) + log1p(IncPerCap) + log1p(Poverty) + log1p(ChildPov) + log1p(Professional) + log1p(Office) + log1p(Construction) + log1p(Production), data=Income)
summary(lm.fit1a)
```

```{r}
#Create best model based on results of best subset selection - log (BIC)
lm.fit1b=lm(log1p(Income)~log1p(Men) + log1p(Women) + log1p(Hispanic) + log1p(White) + log1p(Black) + log1p(Native) + log1p(Asian) + log1p(IncPerCap) + log1p(Poverty) + log1p(Professional) + log1p(Service) + log1p(Production), data=Income)
summary(lm.fit1b)
```

```{r}
#Create best model based on results of best subset selection - log (adjusted R^2)
lm.fit1c=lm(log1p(Income)~log1p(Men) + log1p(Women) + log1p(Hispanic) + log1p(White) + log1p(Black) + log1p(Native) + log1p(Asian) + log1p(Pacific) + log1p(IncPerCap) + log1p(Poverty) + log1p(ChildPov) + log1p(Professional) + log1p(Office) + log1p(Construction) + log1p(Production), data=Income)
summary(lm.fit1c)
```

```{r}
#Cp
par(mfrow=c(2,2))
plot(lm.fit1a, col="green")
```

```{r}
#BIC
par(mfrow=c(2,2))
plot(lm.fit1b, col="green")
```

```{r}
#Adjusted R^2
par(mfrow=c(2,2))
plot(lm.fit1c, col="green")
```

### Step 2: Foward Stepwise Selection ###

```{r}
#Foward Stepwise Selection
regfit.fwd=regsubsets(Income~.,data=subset_data,nvmax=16,method="forward")
reg.summary2 = summary(regfit.fwd)
reg.summary2
```

```{r}
#Create plots for RSS, Adjuted R-squared, Mallow's Cp, and BIC - Foward Stepwise
par(mfrow=c(2,2))
#Plot RSS
plot(reg.summary2$rss,xlab="Number of Variables",ylab="RSS",type="l")

#Plot Adjusted R-squared
plot(reg.summary2$adjr2,xlab="Number of Variables",ylab="Adjusted RSq",type="l")
which.max(reg.summary2$adjr2)
points(15,reg.summary2$adjr2[15], col="blue",cex=2,pch=20)

#Plot Mallow's Cp
plot(reg.summary2$cp,xlab="Number of Variables",ylab="Cp",type='l')
which.min(reg.summary2$cp)
points(15,reg.summary2$cp[15],col="blue",cex=2,pch=20)

#Plot BIC
plot(reg.summary2$bic,xlab="Number of Variables",ylab="BIC",type='l')
which.min(reg.summary2$bic)
points(11,reg.summary2$bic[11],col="blue",cex=2,pch=20)
```

```{r}
#Create best model from forward stepwise selection - log (Cp and Adjusted R^2)
lm.fit2a=lm(log1p(Income)~log1p(Men) + log1p(Women) + log1p(Hispanic) + log1p(White) + log1p(Black) + log1p(Native) + log1p(Asian) + log1p(Pacific) + log1p(IncPerCap) + log1p(Poverty) + log1p(ChildPov) + log1p(Professional) + log1p(Service) + log1p(Office) + log1p(Construction), data=Income)
summary(lm.fit2a)
```

```{r}
#Create best model from forward stepwise selection - log (BIC)
lm.fit2b=lm(log1p(Income)~log1p(Men) + log1p(Women) + log1p(White) + log1p(Black) + log1p(Asian) + log1p(Pacific) + log1p(IncPerCap) + log1p(Poverty) + log1p(Professional) + log1p(Service) + log1p(Construction), data=Income)
summary(lm.fit2b)
```

```{r}
#Cp and adjusted R^2
par(mfrow=c(2,2))
plot(lm.fit2a, col="purple")
```

```{r}
#BIC
par(mfrow=c(2,2))
plot(lm.fit2b, col="purple")
```

### Step 3: Backward Stepwise Selection ###
```{r}
#Backward Stepwise Selection
regfit.bwd=regsubsets(Income~.,data=subset_data,nvmax=16,method="backward")
reg.summary3 = summary(regfit.bwd)
reg.summary3
```

```{r}
#Create plots for RSS, Adjuted R-squared, Mallow's Cp, and BIC - Backward Stepwise
par(mfrow=c(2,2))
#Plot RSS
plot(reg.summary3$rss,xlab="Number of Variables",ylab="RSS",type="l")

#Plot Adjusted R-squared
plot(reg.summary3$adjr2,xlab="Number of Variables",ylab="Adjusted RSq",type="l")
which.max(reg.summary3$adjr2)
points(15,reg.summary3$adjr2[15], col="blue",cex=2,pch=20)

#Plot Mallow's Cp
plot(reg.summary3$cp,xlab="Number of Variables",ylab="Cp",type='l')
which.min(reg.summary3$cp)
points(14,reg.summary3$cp[14],col="blue",cex=2,pch=20)

#Plot BIC
plot(reg.summary3$bic,xlab="Number of Variables",ylab="BIC",type='l')
which.min(reg.summary3$bic)
points(11,reg.summary3$bic[11],col="blue",cex=2,pch=20)
```

```{r}
#Create best model from backward stepwise selection - log (Cp)
lm.fit3a=lm(log1p(Income)~log1p(Men) + log1p(Women) + log1p(Hispanic) + log1p(White) + log1p(Black) + log1p(Native) + log1p(Asian) + log1p(IncPerCap) + log1p(Poverty) + log1p(ChildPov) + log1p(Professional) + log1p(Office) + log1p(Construction) + log1p(Production), data=Income)
summary(lm.fit3a)
```

```{r}
#Create best model from backward stepwise selection - log (BIC)
lm.fit3b=lm(log1p(Income)~log1p(Men) + log1p(Women) + log1p(White) + log1p(Black) + log1p(Asian) + log1p(IncPerCap) + log1p(Poverty) + log1p(Professional) + log1p(Office) + log1p(Construction) + log1p(Production), data=Income)
summary(lm.fit3b)
```

```{r}
#Create best model from backward stepwise selection - log (adjusted R^2)
lm.fit3c=lm(log1p(Income)~log1p(Men) + log1p(Women) + log1p(Hispanic) + log1p(White) + log1p(Black) + log1p(Native) + log1p(Asian) + log1p(Pacific) + log1p(IncPerCap) + log1p(Poverty) + log1p(ChildPov) + log1p(Professional) + log1p(Office) + log1p(Construction) + log1p(Production), data=Income)
summary(lm.fit3c)
```

```{r}
#Cp
par(mfrow=c(2,2))
plot(lm.fit3a, col="darkblue")
```

```{r}
#BIC
par(mfrow=c(2,2))
plot(lm.fit3b, col="darkblue")
```

```{r}
#Adjusted R^2
par(mfrow=c(2,2))
plot(lm.fit3c, col="darkblue")
```

### Step 4: Hybrid Approach ###
```{r}
#seqrep
regfit.hybrid=regsubsets(Income~.,subset_data,nvmax=16,method="seqrep")
reg.summary4 = summary(regfit.hybrid)
reg.summary4
```

```{r}
#Create plots for RSS, Adjuted R-squared, Mallow's Cp, and BIC - Hybrid
par(mfrow=c(2,2))
#Plot RSS
plot(reg.summary4$rss,xlab="Number of Variables",ylab="RSS",type="l")

#Plot Adjusted R-squared
plot(reg.summary4$adjr2,xlab="Number of Variables",ylab="Adjusted RSq",type="l")
which.max(reg.summary4$adjr2)
points(15,reg.summary3$adjr2[15], col="purple",cex=2,pch=20)

#Plot Mallow's Cp
plot(reg.summary4$cp,xlab="Number of Variables",ylab="Cp",type='l')
which.min(reg.summary4$cp)
points(14,reg.summary3$cp[14],col="purple",cex=2,pch=20)

#Plot BIC
plot(reg.summary4$bic,xlab="Number of Variables",ylab="BIC",type='l')
which.min(reg.summary4$bic)
points(12,reg.summary3$bic[12],col="purple",cex=2,pch=20)
```

```{r}
#Create best model from backward stepwise selection - log (Cp)
lm.fit4a=lm(log1p(Income)~log1p(Men) + log1p(Women) + log1p(Hispanic) + log1p(White) + log1p(Black) + log1p(Native) + log1p(Asian) + log1p(IncPerCap) + log1p(Poverty) + log1p(ChildPov) + log1p(Professional) + log1p(Office) + log1p(Construction) + log1p(Production), data=Income)
summary(lm.fit4a)
```


```{r}
#Create best model from backward stepwise selection - log (BIC)
lm.fit4b=lm(log1p(Income)~log1p(Men) + log1p(Women) + log1p(Hispanic) + log1p(White) + log1p(Black) + log1p(Native) + log1p(Asian) + log1p(IncPerCap) + log1p(Poverty) + log1p(Professional) + log1p(Service) + log1p(Construction) + log1p(Production), data=Income)
summary(lm.fit4b)
```

```{r}
#Create best model from backward stepwise selection - log (Adjusted R^2)
lm.fit4c=lm(log1p(Income)~log1p(Men) + log1p(Women) + log1p(Hispanic) + log1p(White) + log1p(Black) + log1p(Native) + log1p(Asian) + log1p(Pacific) + log1p(IncPerCap) + log1p(Poverty) + log1p(Professional) + log1p(Service) + log1p(Construction) + log1p(Production), data=Income)
summary(lm.fit4c)
```


```{r}
#Cp
par(mfrow=c(2,2))
plot(lm.fit4a, col="darkgreen")
```

```{r}
#BIC
par(mfrow=c(2,2))
plot(lm.fit4b, col="darkgreen")
```

```{r}
#Adjusted R^2
par(mfrow=c(2,2))
plot(lm.fit4c, col="darkgreen")
```



